---
title: "manuscript"
author: "Kelly Balfour"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Plants may exhibit specific functional traits in response to local environmental conditions. Previous studies have determined that leaf anatomy and plant size of *Acer* (maple) species can vary depending on the habitat occupied; for example, *Acer* species may exhibit larger and thinner leaves under forest canopy shading [@lei1990]. Nutrient supplementation is also known to influence plant traits, with widespread fertilizer use being common practise in agriculture. Calcium is vital to cell growth, but plant calcium requirements are typically low; therefore, the impact of calcium supplementation on leaf anatomy and plant size could be variable [@burstrom1968]. This data project investigates whether and how calcium availability influences leaf and stem traits in *Acer saccharum* (sugar maple), a culturally, economically, and ecologically important tree species.

## Methods

The dataset used, "hbr_maples" [@juice], was loaded from the R package "lterdatasampler" [@hbrmapl]. Sampling took place during August 2003 and June 2004 at the Hubbard Brook Experimental Forest in Woodstock, New Hampshire, USA [@hbrmapl]. At two watershed sites, reference and calcium-treated, researchers sampled *Acer saccharum* (sugar maple tree) seedlings within transects that were growing 10 steps apart [@hbrmapl]. The height of each seedling (mm) was recorded and the stems were collected; from each stem, two leaf areas (cm2), one leaf dry mass (g), and stem dry mass (g) were measured [@juice].

## Results

Will be added!

## Discussion

This is what I discovered from these data, wow!

## Conclusion

Maple trees are super cool!

## References

```{r}
#Zotero was linked to RStudio; citations added to this file are automatically updated in the "references.bib" file located in the "04_manuscript" folder. 
```

[@burstrom1968]

[@hbrmapl]

[@juice]

[@lei1990]
